---
title: "Problem Set 1, Exercises 2-3"
author: "Abdallah Aboelela, Hussein El Kheshen"
date: "22 October 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Exercise 2, Set Up**
```{r exercise2 setup}
#Importing data
fname = paste("https://raw.githubusercontent.com/jtorcasso/teaching/",
              "master/econ210_fall2017/data/psid_1968.csv", sep="")
df = read.csv(fname)
#Undertanding data
colnames(df)
str(df)

#Defining hourly wage
df$wage <- ifelse(df$inc_labor==0 | df$hours==0, NaN, df$inc_labor/df$hours)
#Working vector
df$D <- ifelse(is.nan(df$wage), 0, 1)
```

**Exercise 2.a**

```{r exercise2.a}
non_missing_data_count <- sum(df$D)
num_observations <- nrow(df)
missing_data_prop <- 1 - non_missing_data_count/num_observations
```

There are `r non_missing_data_count` non-missing data points. The proportion of data that is missing is `r round(missing_data_prop, digits = 3)`.

**Exercise 2.b**

The sample mean for wages is not a good estimate for of the population mean $E[W]$ because it does not take into account the 0 wages of people not working. Instead, the probability limit of the sample mean of wages is $\lim_{n\to\infty} \bar{W}_n = E[W | D = 1]$. To measure $E[W]$ we would need to find information on what currently non-working people's wages *would have been* had they been working.


\pagebreak

**Exercise 2.c**

```{r exercise2.c}
mean_ages_w <- mean(df$age[df$D == 1])
mean_ages_nw <- mean(df$age[df$D == 0])

mean_children_w <- mean(df$hhchildren[df$D == 1])
mean_children_nw <- mean(df$hhchildren[df$D == 0])

mean_edu_w <- mean(df$edu[df$D == 1])
mean_edu_nw <- mean(df$edu[df$D == 0])

mean_fam_inc_w <- mean(df$fam_inc[df$D == 1])
mean_fam_inc_nw <- mean(df$fam_inc[df$D == 0])
```

The average age for a working person is `r round(mean_ages_w, digits=1)`. The average age for a non-working person is `r round(mean_ages_nw, digits=1)`.

The average number of children for a working person is `r round(mean_children_w, digits=1)`. The average number of children for a non-working person is `r round(mean_children_nw, digits=1)`.

The average number of years of education for a working person is `r round(mean_edu_w, digits=1)`. The average number of years of education for a non-working person is `r round(mean_edu_nw, digits=1)`.

The average family income for a working person is `r paste0("$", formatC(as.numeric(mean_fam_inc_w), format="f", digits=2, big.mark=","))`. The average family income for a non-working person is `r paste0("$", formatC(as.numeric(mean_fam_inc_nw), format="f", digits=2, big.mark=","))`.


\pagebreak

**Exercise 3**
```{r exercise3}
df$hhnum = ifelse(df$married==1, df$hhchildren + 2,
                  ifelse(is.na(df$married), NaN, df$hhchildren + 1))
df$poverty = ifelse(df$fam_inc < 12488 & df$hhnum <= 1, 1,
                    ifelse(df$fam_inc < 15877 & df$hhnum <= 2, 1,
                    ifelse(df$fam_inc < 19515 & df$hhnum <= 3, 1,
                    ifelse(df$fam_inc < 25094 & df$hhnum <= 4, 1,
                    ifelse(df$fam_inc < 29714 & df$hhnum <= 5, 1,
                    ifelse(df$fam_inc < 33618 & df$hhnum >= 6, 1,
                    ifelse(is.na(df$fam_inc) | is.na(df$hhnum), NaN, 0)
                    ))))))
```

**Exercise 3.a**

To test whether the poverty rate is greater than 10% we can sum the vector df\$poverty and divide by the length of the same vector. The length of the vector is the number of households measured, and df$poverty is 0 when the family is household is not considered poor.

$H_0$: $\mu$ $\leq$ 10%

$H_1$: $\mu$ $>$ 10%

```{r exercise3.a}
poverty_rate = sum(df$poverty)/length(df$poverty)
```

The poverty rate is `r paste(round(100*poverty_rate, 2), "%", sep="")`, thus we reject the null hypothesis.

**Exercise 3.b**

```{r exercise3.b}
# Defining inputs
alpha <- 0.05
z <- qnorm(1 - alpha)
N <- nrow(df)
effect_size <- 0.01 # Delta c in notes
sd <- sd(df$poverty)

# Calculating power
power <- pnorm(-z + sqrt(N)*effect_size/sd)
```

As shown in class, the power of a test is equal to $\phi$(-$z_{1-\alpha}$ + $\frac{\sqrt{N}*\Delta c}{\sigma})$

The test power with an effect size 0.01 is `r paste(round(100*power, 2), "%", sep="")`. This means that given that the alternative hypothesis is true and we *should* reject the null hypothesis, then we have a `r paste(round(100*power, 2), "%", sep="")` of rejecting the null hypothesis given an $\alpha$ of 0.05 and and an effect size of 0.01.

This is not a particularly strong result and indicates that at the effect size of 0.01 we would be making a significant number of type II errors.

**Exercise 3.c**

$H_0$: $\mu$ = 10%

$H_1$: $\mu$ $\neq$ 10%

```{r exercise3.c}
# Defining inputs
mean <- mean(df$poverty)
test <- 0.1
N <- length(df$poverty)
sigma <- sd(df$poverty)

# Calculating t
t <- (mean - test)*sqrt(N) / sigma

# Calculating p-value
p_value = 2*(1-pnorm(abs(t)))
```

With a p-value of `r p_value`, we can safely reject the null hypothesis at the 0.05 and 0.01 significance levels.

**Exercise 3.d**
```{r exercise3.d}
# Defining inputs
mean <- mean(df$poverty)
alpha <- 0.05
z <- qnorm(1-alpha/2)
N <- length(df$poverty)
var <- var(df$poverty)

# Calculating bounds of interval
lower_bound <- mean - z * sqrt(var/N)
upper_bound <- mean + z * sqrt(var/N)
```

Our 95% confidence interval of the mean is `r sprintf("(%s, %s)",round(lower_bound, 4),round(upper_bound, 4))`.

**Exercise 3.e**

We would like to determine whether it is reasonable to assume that the proportion of the surveyed population who are married and have children is representative of the population of the U.S. who are married and have children in 2011 as shown by Statista.

(Source: https://www.statista.com/statistics/242074/percentages-of-us-family-households-with-children-by-type/)

Our test-statistic is a binary variable equal to 1 if the household is married and has children and 0 if the household is married and has no children.

$H_0$: $\mu$ = 0.4

$H_1$: $\mu \neq$ 0.4

```{r exercise3.e}
# Importing 2011 data
fname <- paste("https://raw.githubusercontent.com/jtorcasso/teaching/",
"master/econ210_fall2017/data/psid_2011.csv", sep="")
df2011 <- read.csv(fname)

# Building binary variable
df2011$ch <- ifelse(df2011$married == 1 & df2011$hhchildren > 0, 1, 
                    ifelse(df2011$married == 1 & df2011$hhchildren == 0, 0,
                    ifelse(df2011$married == 0, NA, NA)))

ch <- df2011$ch[!is.na(df2011$ch)]

# Defining variables
mean <- mean(ch)
test <- 0.4
N <- length(ch)
sigma <- sd(ch)

# Calculating t
t <- (mean - test)*sqrt(N) / sigma

# Calculating p-value
p_value <- 2*(1-pnorm(abs(t)))

# Confidence interval
z <- qnorm(1-alpha/2)
N <- length(ch)
var <- var(ch)

# Calculating bounds of interval
lower_bound <- mean - z * sqrt(var/N)
upper_bound <- mean + z * sqrt(var/N)
```

By running similar tests to Exercises 3.c and 3.d, we find that we have a p-value of `r p_value` and a confidence interval of `r sprintf("(%s, %s)", round(lower_bound, 4), round(upper_bound, 4))`. 

This leads us to conclude that the likelihood that married couples in the sample is not representative of the likelihood of married couples in the U.S. generally to have children. This sample is much more likely to have children when married than the average married couple in the U.S.

From a practical perspective, we believe that our t-score of `r t` leads to absurdly low values for the p-value which R rounds down to 0.